# Regression

## Linear Regression Models

Linear Regression or Linear Model (LM) using happiness data from the World Health Organization, this section is brought to you by **RLadies Freiburg** Divya and Elisa. The happiness data can be found [here](https://raw.githubusercontent.com/rladies/meetup-presentations_freiburg/master/2019-10-Modelling%201%20-%20LM%2C%20LMM%20and%20plots/Happiness_2016.csv). 

```{r echo=FALSE}
# load in the data (the dataset is read in as raw)
happiness_2016 = read.csv('https://raw.githubusercontent.com/rladies/meetup-presentations_freiburg/master/2019-10-Modelling%201%20-%20LM%2C%20LMM%20and%20plots/Happiness_2016.csv')

knitr::kable(head(happiness_2016) ) 
```



The question is, *what parameters are there? What questions can we address using these factors.?* *Does trust in government a variable?* 

```{r pressure, echo=TRUE}
plot(happiness_2016$Happiness.Score, happiness_2016$Trust..Government.Corruption.)
```

### try some models

There are 3 questions we could ask:
- Does Corruption within the government predict happiness of its citizens?
- How can we add Freedom as a predictor to the model?
- What is the difference between adding factors and looking at their interactions?

```{r}
## ANSWERS
#fit our regression model
RegressionModel <- lm(Happiness.Score ~ Trust..Government.Corruption., # regression formula
              data=happiness_2016) # data set

# Summarize and print the results
summary(RegressionModel) # show regression coefficients table


RegressionModel2 <- lm(Happiness.Score ~ Trust..Government.Corruption. + Freedom,
                      data=happiness_2016) 
summary(RegressionModel2) 


RegressionModel3 <- lm(Happiness.Score ~ Trust..Government.Corruption. * Freedom, 
                       data=happiness_2016) 

summary(RegressionModel3)
```

There are a number of different Parametric and non-parametric tests we can try in the same format. Instead of lm, we could use t.test(), aov(), wilcox.test(), kruskal.test().

Is there a categorical variable we can use to run an ANOVA on?


*How does the happiness quotient differ based on regions?*

```{r}
ANOVA <- aov(Happiness.Score ~ Region, 
             data=happiness_2016) 
summary(ANOVA) 
TukeyHSD(ANOVA)
```


### independence of the data

Let's think about the data and pose some questions. 


A) Do you think that countries in the same region tend to be more similar to each other? 
If your answer is yes, then the countries are not really independent and identically distributed data. This could be a problem with statistical models. So you have three options:

- 1. Ignore the problem, or argue why you think your data IS independent (you will no be the only one)
- 2. Do not ignore it, take the average of each region and then make a model using this average. Yes, what you think is correct, you loose a lot of data!
- 3. Do one linear model for each region...data costly (e.g. Australia and North America have only two data points)
- 4. Try a linear mixed model


B) How variable is the happiness between regions?


C) How does happiness depend on other predictors such as health, economy and generosity?


### visualization of the data

We can do some plots to see how data look like.
In this plot we visualize how economy affects happiness in every region. We could do one plot per predictor. 

```{r}
library(dplyr)
library(tidyr)
Happiness_2016 <- happiness_2016 %>% group_by(Region) %>% mutate(mean.reg = mean(Happiness.Score)) %>%  ungroup()
library(ggplot2)
ggplot(Happiness_2016) + 
  aes(x = Economy..GDP.per.Capita., y = Happiness.Score) + 
  stat_smooth(method = "lm", se = FALSE) +
  # Put the points on top of lines
  geom_point() +
  facet_wrap("Region") +
  labs(x = "Economy", y = "Happyness") +
  geom_hline(aes(yintercept = mean.reg), colour='red', lty="dotted")
```


Given what we see in the data, we can try two different models: 

```{r}
library(lme4)
happy.mixed.model <-  lmer(Happiness.Score ~ Economy..GDP.per.Capita. + Health..Life.Expectancy.+ Generosity + (1|Region), data = happiness_2016)

summary(happy.mixed.model)

happy.mixed.model.2 <-  lmer(Happiness.Score ~ Economy..GDP.per.Capita. + Health..Life.Expectancy.+ Generosity + (Economy..GDP.per.Capita.|Region), 
                             data = happiness_2016)

summary(happy.mixed.model)
summary(happy.mixed.model.2)
```

We do not have p-values here hummm!

```{r}
library(lmerTest)
# Test significance of random effects
ranova(happy.mixed.model.2)

#Test significance of fixed effects
ml.happy.mixed.model <- update(happy.mixed.model, REML = FALSE) # this changes the algorithm used to fit the model

#Finally we can test significance
anova(as_lmerModLmerTest(ml.happy.mixed.model))
```



### Ploting the data in a map

```{r}
require(rnaturalearth)
require(rnaturalearthdata)
world <- ne_countries(scale = "medium", returnclass = "sf") # this is another function to get polygons of countries. 
```


```{r}
library(ggplot2)
ggplot(data = world) +
    theme_bw()+ 
    geom_sf() + 
    xlab("Longitude") + ylab("Latitude") + 
    ggtitle("World map", subtitle = paste0("(", dim(world)[1], " countries)"))
```


```{}
#both df woth the same name of the varaible we will use to join
colnames(happiness_2016)[1] <- "name"
Happiness_GEO <- left_join(world, happiness_2016, by="name")
```
```{}
ggplot(data = Happiness_GEO) +
    geom_sf(aes(fill = Happiness.Score )) +
    scale_fill_viridis_c(option = "plasma") +  # this allows you to choose different colour scale
    ggtitle("World Happiness Studies")
```

#### Where is the model working better?

```{}
happy.predictions <- predict(happy.mixed.model)
#Expected - Observed
happy.residuasl <-  (happiness_2016$Happiness.Score - happy.predictions) # Obaserved - predicted
Happiness_2016$Model.Residuals <- happy.residuasl
Happiness_GEO <- left_join(world, Happiness_2016, by="name")
```

```{}
ggplot(data = Happiness_GEO) +
    geom_sf(aes(fill = Model.Residuals )) +
    scale_fill_viridis_c(option = "plasma") +  # this allows you to choose different colour scale
    ggtitle("World Happiness Studies")
```





